## Machine Learning Model Building

Agenda:

1.  Loading Necessary Libraries.

2.  Load the dataset.

3.  Selecting of relevant variables for Abstraction and Generalization.

4.  Feature Engineering (for accurate modelling)

5.  Datasets Splitting (Training and Testing).

6.  Model Building (with Linear Regression, Random Forest and Decision Tree algorithms

```{r}
#Load necessary libraries
library(tidyverse)
library(caret)
library(randomForest)
library(corrplot)
```

```{r}
#Load our dataset into Global Env
diamonds_df <- read_csv("datasets/diamonds.csv")
```

```{r}
diamonds_df[, c("carat", "cut", "color", "clarity", "depth", "table", "price")]
```

## Correlation Matrix

```{r}
#Correlation MAtrix
correlation_matrix <- cor (
   diamonds_df[, c("carat", "depth", "table", "x", "y", "z", "price")]
)

```

```{r}
print("Correlation MAtrix:")
print(correlation_matrix)
```

```{r}
#Visualize the correlation matrix
corrplot(correlation_matrix,method = "color", addCoef.col = "black")
```

```{r}
#Feature engineering
diamonds_selected <- diamonds_df[, c("carat", "cut", "color", "clarity", "x", "y", "z", "price")]
diamonds_selected
```

## Splitting the dataset into training and testing

```{r}
#splitting the dataset into training and tests (70/30 ratio)
set.seed(56)
train_index <- createDataPartition(diamonds_selected$price, p = 0.7, list = FALSE,)
train_data <- diamonds_selected[train_index, ]
test_data <- diamonds_selected[- train_index, ]
```

```{r}
test_data
```

### Model Building Using the Linear Regression Algorithm

Linear regression is a statistical method used to model a relationship between the dependent variable (target)

```{r}
#linear Regression
linear_model <- lm(price ~ ., data = train_data)
```

```{r}
#Evaluate linear Regression model
linear_predictions_train <- predict(linear_model, newdata = train_data)
linear_mse_train <- mean((train_data$price - linear_predictions_train)^2)
cat("\nLinear Regression - Mean Squared Error (Train):", linear_mse_train, "\n")

linear_predictions_test <- predict(linear_model, newdata = test_data)
linear_mse_test <- mean((test_data$price - linear_predictions_test)^2)
cat("Linear Regression - Mean Squared Error (Test):", linear_mse_test, "\n")
```
